---
title: "Chapman-Richards Curve Parameter Estimation"
subtitle: "Pixel-level AGC growth curve fitting for natural forest regeneration"
author: "Nathaniel Robinson / adapted by Serge Wiltshire"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: show
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
rm(list = ls())
```

---

## Overview

This notebook fits a **Chapman-Richards (CR)** growth curve to pixel-level above-ground
carbon (AGC) estimates from random forest outputs. For each pixel, the model is:

$$
\text{AGC}(t) = A \cdot \left(1 - B \cdot e^{-K \cdot t}\right)^{\frac{1}{1-m}}, \quad m = \tfrac{2}{3}
$$

which simplifies to:

$$
\text{AGC}(t) = A \cdot \left(1 - B \cdot e^{-K \cdot t}\right)^{3}
$$

**Parameters estimated per pixel:**

| Parameter | Meaning |
|-----------|---------|
| A | Asymptote — maximum potential AGC |
| K | Growth rate |
| B | Shape parameter |

**Inputs** are GEE-exported tiles (one per 62°×62° region) at ~1 km resolution:

- `data/inputs/{scenario}/age_{AGE}_{TILE_ID}.tif` — 20-band rasters (random forest
  realizations `sd_001`–`sd_020`) at 20 age steps (5, 10, …, 100 years)
- **future** and **hist** scenarios are run separately by changing `scenario` below.

**Outputs** are written to `data/outputs/{scenario}/` in subdirectories `A/`, `K/`, `B/`,
`A_error/`, `K_error/`, `B_error/`, `convergence/`.

> **Computational note:** Each tile is 7 424 × 7 424 pixels with 400 bands stacked
> (20 ages × 20 realizations). Processing is parallelized at the tile level
> (`foreach` / `doParallel`). Fitting uses `terra::app()` so that rasters are
> read and written in memory-efficient chunks rather than loaded all at once.
> Expect long runtimes on full data — size the core count to available RAM
> (≥16 GB recommended; 24+ cores optimal).

---

## 1. Package Setup

```{r packages}
required_packages <- c(
  "terra",
  "tidyverse",
  "minpack.lm",
  "parallel",
  "doParallel",
  "foreach",
  "tictoc"
)

new_pkgs <- required_packages[
  !(required_packages %in% installed.packages()[, "Package"])
]
if (length(new_pkgs) > 0) {
  install.packages(new_pkgs, dep = TRUE)
}

invisible(lapply(required_packages, library, character.only = TRUE))
```

---

## 2. Configuration

Adjust these parameters before each run.

```{r config}
# ── Smoke test toggle ─────────────────────────────────────────────────────────
# TRUE  → run on one tile cropped to test_crop_n × test_crop_n pixels.
#         Completes in minutes on a laptop. Use this to verify the fitting
#         pipeline and inspect example curves before a full production run.
# FALSE → full production run over all tiles.
test_mode    <- FALSE
test_crop_n  <- 150   # pixels per side when test_mode = TRUE

# ── Scenario ──────────────────────────────────────────────────────────────────
# Set to "future" or "hist" — controls which input directory is used.
# To process both scenarios back-to-back, see Section 6.
scenario <- "hist"

# ── Paths ─────────────────────────────────────────────────────────────────────
# Notebook-relative root of the project (two levels up from cr_calc/).
# On SageMaker, set this explicitly to the mounted S3/EFS path, e.g.:
#   project_root <- "/home/ec2-user/SageMaker/nat_for_regen_c_accumulation_sw"
project_root <- normalizePath(file.path(dirname(rstudioapi::getActiveDocumentContext()$path),
                                         ".."))

input_dir  <- file.path(project_root, "data", "inputs",  scenario)
output_dir <- file.path(project_root, "data", "outputs", scenario)
log_file   <- file.path(output_dir, sprintf("progress_%s.log", scenario))

# ── Age sequence ──────────────────────────────────────────────────────────────
# Ages (years) matching the input file naming convention (age_005, age_010, …)
ages    <- seq(5, 100, by = 5)   # 20 time steps
n_bands <- 20                    # realizations per age file (sd_001 – sd_020)

# Predictor vector for NLS: each age is repeated once per realization band,
# producing a length-400 vector aligned to the stacked 400-band raster.
age_vector <- rep(ages, each = n_bands)

# ── Parallelisation strategy ───────────────────────────────────────────────────
# Two modes:
#
#   "tile"  — one worker per tile (18 tiles max). Best when n_cores ≤ 18.
#             Uses socket or fork clusters via doParallel/foreach.
#
#   "pixel" — tiles run sequentially; terra::app() parallelises across
#             pixels within each tile using all available cores.
#             Best on large instances (SageMaker ≥ 18 vCPUs) because it
#             scales to any core count. On Linux, terra uses fork workers
#             (zero-copy, no serialization overhead).
#
# Auto-select: use "pixel" when n_cores > n_tiles, else "tile".
# NOTE: terra::app(cores=n) does not parallelise in the SageMaker Posit
# environment (confirmed via /proc/loadavg). Force "tile" strategy, which
# uses foreach socket workers — one tile per worker, 18 tiles in parallel.
n_tiles  <- 18L   # total tiles per scenario
n_cores  <- max(1L, detectCores() - 1L)
parallel_strategy <- "tile"

# Each tile is split into horizontal strips so that total work units stays
# under R's 128-connection limit while maximising core utilisation.
n_max_workers     <- 80L
n_strips_per_tile <- max(1L, floor(n_max_workers / n_tiles))

cat(sprintf("Cores available: %d | Strategy: %s | Strips/tile: %d | Work units: %d\n",
            n_cores, parallel_strategy, n_strips_per_tile,
            n_tiles * n_strips_per_tile))

# ── Memory ────────────────────────────────────────────────────────────────────
# Maximum RAM (GB) terra may use per process when chunking raster I/O.
# - "tile" mode:  budget = total_RAM / n_cores  (each worker reads one tile)
# - "pixel" mode: budget = ~total_RAM * 0.75    (one tile at a time, all cores)
# SageMaker ml.c5.24xlarge (192 GB): terra_mem_gb <- 120 in "pixel" mode.
terra_mem_gb <- 120
terraOptions(memmax = terra_mem_gb)

# ── S3 output ─────────────────────────────────────────────────────────────────
# TRUE  → after all tiles complete, sync local outputs to S3 (Section 9).
#          Requires rclone configured (see setup/fetch_input_data.Rmd).
# FALSE → outputs stay local only (laptop / smoke test).
save_to_s3 <- TRUE

s3_output_dir <- sprintf(
  "s3://sagemaker-gst-stage.sharing/serge-wiltshire/nat_for_regen_c_accumulation_data/%s/outputs",
  scenario
)
```

---

## 3. Function Definitions

### 3.1 Chapman-Richards model

```{r fn_cr}
#' Chapman-Richards growth model (m fixed at 2/3)
#'
#' @param t  Numeric vector of ages (years).
#' @param A  Asymptote parameter.
#' @param K  Growth rate parameter.
#' @param B  Shape parameter.
#' @return   Numeric vector of predicted AGC values.
chapman_richards <- function(t, A, K, B) {
  A * (1 - B * exp(-K * t))^3
}
```

### 3.2 Single NLS attempt

```{r fn_fit_nls}
#' Attempt one NLS fit of the CR curve.
#'
#' @param pix_y   Numeric response vector (AGC values for one pixel,
#'                length = n_ages * n_bands = 400).
#' @param age_v   Numeric predictor vector of the same length as pix_y.
#' @param a_start Starting value for A (typically max_pot for this pixel).
#' @param k_start Starting value for K.
#' @param b_start Starting value for B.
#' @return Named numeric vector (A, K, B, A_err, K_err, B_err) on success,
#'         or NULL on failure.
attempt_nls <- function(pix_y, age_v,
                        a_start,
                        k_start = 0.05,
                        b_start = 0.75) {
  lower <- c(a = a_start * 0.9, k = 0.01, b = 0.2)
  upper <- c(a = a_start * 1.1, k = 0.10, b = 1.0)

  fit <- try(
    minpack.lm::nlsLM(
      pix_y ~ a * (1 - b * exp(-k * age_v))^3,
      start   = list(a = a_start, k = k_start, b = b_start),
      lower   = lower,
      upper   = upper,
      control = minpack.lm::nls.lm.control(maxiter = 500)
    ),
    silent = TRUE
  )

  if (inherits(fit, "try-error")) return(NULL)

  cf <- coef(fit)
  se <- tryCatch(
    summary(fit)$coefficients[, "Std. Error"],
    error = function(e) rep(NA_real_, 3)
  )

  c(A     = unname(cf["a"]),
    K     = unname(cf["k"]),
    B     = unname(cf["b"]),
    A_err = se[1],
    K_err = se[2],
    B_err = se[3])
}
```

### 3.3 Multi-iteration pixel fitting

```{r fn_fit_pixel}
#' Fit CR curve for a single pixel with up to 3 NLS iterations.
#'
#' Iteration strategy:
#'   1. Start from max_pot + default k/b.
#'   2. If iteration 1 did not fully converge (stopCode ≠ 0),
#'      re-start from iteration 1 coefficients.
#'   3. If iteration 2 still did not converge, re-start from iteration 2
#'      coefficients.
#'   4. If all three fail, return NA with convergence code 4.
#'
#' Convergence codes: 1 = converged on 1st try, 2 = 2nd, 3 = 3rd, 4 = failed.
#'
#' @param pix_y   Numeric response vector (length = n_ages * n_bands).
#' @param age_v   Numeric predictor vector (same length as pix_y).
#' @param max_pot Maximum potential AGC for this pixel (used to anchor A).
#' @return Named numeric vector of length 7:
#'         A, K, B, A_err, K_err, B_err, convergence.
fit_cr_pixel <- function(pix_y, age_v, max_pot) {
  na_result  <- c(A = NA_real_, K = NA_real_, B = NA_real_,
                  A_err = NA_real_, K_err = NA_real_, B_err = NA_real_,
                  convergence = NA_real_)
  zero_result <- c(A = 0, K = 0, B = 0,
                   A_err = 0, K_err = 0, B_err = 0,
                   convergence = 0)

  if (all(is.na(pix_y)))              return(na_result)
  if (any(is.na(pix_y)))             return(na_result)
  if (is.na(max_pot) | max_pot <= 0) return(zero_result)

  # Helper: check whether nls converged (stopCode == 0 for "port" algorithm)
  converged <- function(fit) {
    tryCatch(
      isTRUE(fit$convInfo$stopCode == 0),
      error = function(e) TRUE  # assume ok if info unavailable
    )
  }

  # ── Iteration 1 ─────────────────────────────────────────────────────────────
  r1 <- attempt_nls(pix_y, age_v, a_start = max_pot)
  if (is.null(r1)) return(c(na_result[1:6], convergence = 4))

  # Retrieve the fit object again briefly just to check stopCode
  # (attempt_nls already ran it; we skip re-running here and use warnOnly
  #  heuristic: if A/K/B are within bounds, treat as converged enough)
  if (r1["A"] >= max_pot * 0.9 && r1["A"] <= max_pot * 1.1) {
    return(c(r1, convergence = 1))
  }

  # ── Iteration 2: re-start from iteration 1 result ───────────────────────────
  r2 <- attempt_nls(pix_y, age_v,
                    a_start = r1["A"],
                    k_start = r1["K"],
                    b_start = r1["B"])
  if (is.null(r2)) return(c(r1, convergence = 4))

  if (r2["A"] >= r1["A"] * 0.9 && r2["A"] <= r1["A"] * 1.1) {
    return(c(r2, convergence = 2))
  }

  # ── Iteration 3: re-start from iteration 2 result ───────────────────────────
  r3 <- attempt_nls(pix_y, age_v,
                    a_start = r2["A"],
                    k_start = r2["K"],
                    b_start = r2["B"])
  if (is.null(r3)) return(c(r2, convergence = 4))

  c(r3, convergence = 3)
}
```

### 3.4 Tile-level CR calculation

```{r fn_calc_tile}
#' Fit CR curves for all pixels in a single tile.
#'
#' Stacks the 400-band raster (20 ages × 20 realizations) for the given tile,
#' then applies pixel-wise NLS fitting via terra::app() — raster chunks are
#' read sequentially so the full tile never lives in RAM at once.
#' Outputs are written as compressed single-band GeoTIFFs.
#'
#' @param tile_id   Character tile identifier extracted from filenames
#'                  (e.g. "17817146248550000000000-0000000000").
#' @param input_dir Full path to the scenario input directory.
#' @param output_dir Full path to the scenario output directory.
#' @param ages      Numeric vector of age steps (default seq(5, 100, by=5)).
#' @param n_bands   Number of realization bands per age file (default 20).
#' @param log_file  Path to log file for progress messages.
#' @param app_cores   Integer number of cores to pass to terra::app().
#'                    Use n_cores in "pixel" strategy; 1L in "tile" strategy.
#' @param test_crop_n  Integer. If > 0, auto-locate valid data and crop to a
#'                     test_crop_n x test_crop_n pixel window. Smoke test only.
#' @param crop_ext     SpatExtent or NULL. If provided, crop the stacked raster
#'                     to this extent before fitting. Used for strip-based
#'                     parallelism. Takes precedence over test_crop_n.
#' @param strip_suffix Character suffix for output filenames (e.g. "_s01").
#' @return Invisibly returns the tile_id on success.
calc_cr_tile <- function(tile_id,
                         input_dir,
                         output_dir,
                         ages         = seq(5, 100, by = 5),
                         n_bands      = 20,
                         log_file     = NULL,
                         app_cores    = 1L,
                         test_crop_n  = 0L,
                         crop_ext     = NULL,
                         strip_suffix = "") {

  age_v <- rep(ages, each = n_bands)  # predictor vector, length 400

  # ── Build file list (sorted by age) ─────────────────────────────────────────
  age_files <- file.path(input_dir,
                         sprintf("age_%03d_%s.tif", ages, tile_id))

  missing_files <- age_files[!file.exists(age_files)]
  if (length(missing_files) > 0) {
    msg <- sprintf("[%s] SKIPPED — missing files:\n  %s\n",
                   tile_id, paste(missing_files, collapse = "\n  "))
    message(msg)
    if (!is.null(log_file)) cat(msg, file = log_file, append = TRUE)
    return(invisible(NULL))
  }

  # ── Stack into a 400-band SpatRaster (lazy — not read into RAM yet) ──────────
  agc_stack <- rast(age_files)  # terra stacks bands in file order (age order)

  # ── Optional: crop to strip extent or smoke-test window ──────────────────────
  if (!is.null(crop_ext)) {
    # crop_ext arrives as a numeric vector (serialization-safe); reconstruct
    agc_stack <- crop(agc_stack, ext(crop_ext))
  } else if (test_crop_n > 0) {
    # Locate valid data: aggregate one layer coarsely to find a non-zero region,
    # then centre the crop window on the densest block.
    probe   <- rast(age_files[length(age_files)])[[1]]  # last age, first band
    agg_fct <- max(1L, floor(min(nrow(probe), ncol(probe)) / 75L))
    probe_agg <- aggregate(probe, fact = agg_fct, fun = "max", na.rm = TRUE)
    agg_vals  <- values(probe_agg, mat = FALSE)
    best_cell <- which.max(replace(agg_vals, is.na(agg_vals), 0))

    if (length(best_cell) == 0 || max(agg_vals, na.rm = TRUE) == 0) {
      msg_crop <- sprintf(
        "[%s] TEST MODE: no valid data found — cropping top-left %d×%d px\n",
        tile_id, test_crop_n, test_crop_n)
      e  <- ext(agc_stack)
      rx <- res(agc_stack)[1]; ry <- res(agc_stack)[2]
      crop_ext <- ext(e[1], e[1] + test_crop_n*rx,
                      e[4] - test_crop_n*ry, e[4])
    } else {
      best_row <- rowFromCell(probe_agg, best_cell)
      best_col <- colFromCell(probe_agg, best_cell)
      # Convert aggregated row/col back to original pixel centre
      mid_row  <- (best_row - 0.5) * agg_fct
      mid_col  <- (best_col - 0.5) * agg_fct
      e  <- ext(agc_stack)
      rx <- res(agc_stack)[1]; ry <- res(agc_stack)[2]
      half <- test_crop_n / 2
      x_mid <- e[1] + (mid_col - 0.5) * rx
      y_mid <- e[4] - (mid_row - 0.5) * ry
      crop_ext <- ext(x_mid - half*rx, x_mid + half*rx,
                      y_mid - half*ry, y_mid + half*ry)
      msg_crop <- sprintf(
        "[%s] TEST MODE: cropped %d×%d px centred on lon=%.2f lat=%.2f\n",
        tile_id, test_crop_n, test_crop_n, x_mid, y_mid)
    }

    agc_stack <- crop(agc_stack, crop_ext)
    message(msg_crop)
    if (!is.null(log_file)) cat(msg_crop, file = log_file, append = TRUE)
  }  # end test_crop_n

  # ── Define the pixel-wise fitting function (fully self-contained closure) ────
  #
  # terra::app() calls this function once per pixel, passing x = numeric vector
  # of length nlyr(agc_stack) = 400 (one value per band).
  # The function must return a fixed-length numeric vector; here length 7.
  #
  # All helper functions are defined INSIDE the closure via local() so they are
  # captured in the function environment and available in any worker context
  # (fork or socket) without relying on name lookup in the global environment.
  cr_app_func <- local({

    av <- age_v   # capture predictor vector

    .attempt_nls <- function(pix_y, age_v, a_start, k_start = 0.05, b_start = 0.75) {
      lower <- c(a = a_start * 0.9, k = 0.01, b = 0.2)
      upper <- c(a = a_start * 1.1, k = 0.10, b = 1.0)
      fit <- try(minpack.lm::nlsLM(
        pix_y ~ a * (1 - b * exp(-k * age_v))^3,
        start   = list(a = a_start, k = k_start, b = b_start),
        lower   = lower,
        upper   = upper,
        control = minpack.lm::nls.lm.control(maxiter = 500)
      ), silent = TRUE)
      if (inherits(fit, "try-error")) return(NULL)
      cf <- coef(fit)
      se <- tryCatch(summary(fit)$coefficients[, "Std. Error"],
                     error = function(e) rep(NA_real_, 3))
      c(A = unname(cf["a"]), K = unname(cf["k"]), B = unname(cf["b"]),
        A_err = se[1], K_err = se[2], B_err = se[3])
    }

    .fit_cr_pixel <- function(pix_y, age_v, max_pot) {
      na_r <- c(A = NA_real_, K = NA_real_, B = NA_real_,
                A_err = NA_real_, K_err = NA_real_, B_err = NA_real_,
                convergence = NA_real_)
      if (all(is.na(pix_y)) | any(is.na(pix_y))) return(na_r)
      if (is.na(max_pot) | max_pot <= 0)
        return(c(A=0, K=0, B=0, A_err=0, K_err=0, B_err=0, convergence=0))
      r1 <- .attempt_nls(pix_y, age_v, a_start = max_pot)
      if (is.null(r1)) return(c(na_r[1:6], convergence = 4))
      if (r1["A"] >= max_pot * 0.9 & r1["A"] <= max_pot * 1.1) return(c(r1, convergence = 1))
      r2 <- .attempt_nls(pix_y, age_v, a_start = r1["A"], k_start = r1["K"], b_start = r1["B"])
      if (is.null(r2)) return(c(r1, convergence = 4))
      if (r2["A"] >= r1["A"] * 0.9 & r2["A"] <= r1["A"] * 1.1) return(c(r2, convergence = 2))
      r3 <- .attempt_nls(pix_y, age_v, a_start = r2["A"], k_start = r2["K"], b_start = r2["B"])
      if (is.null(r3)) return(c(r2, convergence = 4))
      c(r3, convergence = 3)
    }

    function(x) {
      .fit_cr_pixel(pix_y = x, age_v = av, max_pot = max(x, na.rm = TRUE))
    }
  })

  # ── Apply pixel-wise (terra handles chunked I/O) ─────────────────────────────
  label <- paste0(tile_id, strip_suffix)
  msg_start <- sprintf("[%s] Starting CR fit — %s\n",
                       label, format(Sys.time(), "%H:%M:%S"))
  if (!is.null(log_file)) cat(msg_start, file = log_file, append = TRUE)

  # app_cores > 1 uses terra's built-in parallelism (fork on Linux, socket on Windows)
  result <- app(agc_stack, fun = cr_app_func, cores = app_cores)
  names(result) <- c("A", "K", "B", "A_error", "K_error", "B_error", "convergence")

  # ── Write one GeoTIFF per parameter ─────────────────────────────────────────
  param_dirs <- file.path(output_dir, names(result))
  invisible(lapply(param_dirs, dir.create, recursive = TRUE, showWarnings = FALSE))

  for (param in names(result)) {
    out_path <- file.path(output_dir, param,
                          sprintf("%s_%s%s.tif", param, tile_id, strip_suffix))
    writeRaster(result[[param]], out_path,
                gdal = "COMPRESS=DEFLATE", overwrite = TRUE)
  }

  msg_done <- sprintf("[%s] Complete — %s\n",
                      label, format(Sys.time(), "%H:%M:%S"))
  message(msg_done)
  if (!is.null(log_file)) cat(msg_done, file = log_file, append = TRUE)

  invisible(tile_id)
}
```

---

## 4. File Discovery

```{r discovery}
# ── List available input tiles ────────────────────────────────────────────────
all_files <- list.files(input_dir, pattern = "\\.tif$")

# Extract tile IDs: everything after "age_XXX_" and before ".tif"
all_tile_ids <- unique(sub("^age_\\d+_(.+)\\.tif$", "\\1", all_files))

cat(sprintf("Scenario: %s\n", scenario))
cat(sprintf("Input directory: %s\n", input_dir))
cat(sprintf("Tiles found: %d\n", length(all_tile_ids)))
cat(sprintf("Ages found: %s\n",
            paste(sort(unique(as.integer(
              sub("^age_(\\d+)_.*$", "\\1", all_files)))), collapse = ", ")))
```

```{r work_units}
if (test_mode) {

  # ── Smoke test: one tile, auto-locate crop, no strip splitting ──────────────
  work_units <- list(list(tile_id = all_tile_ids[1],
                          crop_ext = NULL,
                          strip_suffix = ""))
  cat(sprintf("[TEST MODE] 1 work unit (1 tile × %d×%d px crop)\n",
              test_crop_n, test_crop_n))

} else {

  # ── Production: split each tile into horizontal strips ──────────────────────
  # This creates n_tiles × n_strips_per_tile work units so we can saturate all
  # available cores via foreach, without relying on terra::app(cores=n).
  work_units <- lapply(all_tile_ids, function(tid) {
    # Read extent and resolution from one input file (lazy — no pixel data read)
    probe <- rast(file.path(input_dir, sprintf("age_005_%s.tif", tid)))
    e     <- ext(probe)
    ry    <- res(probe)[2]
    nr    <- nrow(probe)
    rows_per_strip <- ceiling(nr / n_strips_per_tile)

    lapply(seq_len(n_strips_per_tile), function(s) {
      row_start <- (s - 1L) * rows_per_strip + 1L
      row_end   <- min(s * rows_per_strip, nr)
      y_top     <- e[4] - (row_start - 1L) * ry
      y_bot     <- e[4] - row_end * ry
      # Store extent as plain numeric vector (not SpatExtent) so it
      # survives serialization to socket workers. Reconstructed in
      # calc_cr_tile via terra::ext().
      list(tile_id      = tid,
           crop_ext     = c(e[1], e[2], y_bot, y_top),
           strip_suffix = sprintf("_s%02d", s))
    })
  }) |> unlist(recursive = FALSE)

  # ── Skip already-completed work units ───────────────────────────────────────
  completed_dir <- file.path(output_dir, "A")
  if (dir.exists(completed_dir)) {
    completed_files <- list.files(completed_dir, pattern = "\\.tif$")
  } else {
    completed_files <- character(0)
  }
  is_done <- vapply(work_units, function(wu) {
    fname <- sprintf("A_%s%s.tif", wu$tile_id, wu$strip_suffix)
    fname %in% completed_files
  }, logical(1))
  work_units <- work_units[!is_done]

  cat(sprintf("Work units: %d remaining (%d already done)\n",
              length(work_units), sum(is_done)))
}
```

---

## 5. Output Directory Setup

```{r output_dirs}
param_names <- c("A", "K", "B", "A_error", "K_error", "B_error", "convergence")
param_dirs  <- file.path(output_dir, param_names)
invisible(lapply(param_dirs, dir.create, recursive = TRUE, showWarnings = FALSE))

file.create(log_file, showWarnings = FALSE)
cat(sprintf("Run started: %s\nScenario: %s\nWork units: %d\n",
            format(Sys.time()), scenario, length(work_units)),
    file = log_file, append = TRUE)

cat("Output directories ready under:", output_dir, "\n")
```

---

## 6. CR Parameter Fitting

Each work unit (tile × strip) is dispatched to a socket-cluster worker via
`foreach %dopar%`. With 18 tiles × 5 strips/tile = 90 work units, this keeps
~90 cores busy without relying on `terra::app()` internal parallelism.

To process both scenarios back-to-back, change `scenario` in Section 2 and
re-run from Section 2 onward.

```{r fit_parallel}
tic(sprintf("CR fitting — %s (%d work units)", scenario, length(work_units)))

n_workers <- min(n_cores, length(work_units))
cat(sprintf("Running %d work units with %d forked workers\n",
            length(work_units), n_workers))

# mclapply uses fork on Linux — child processes inherit the parent's memory
# directly (no serialization), which avoids the "external pointer is not valid"
# error that occurs when foreach socket clusters try to serialize terra objects.
# Per-worker memory: divide available RAM across workers, leave 20% for OS
per_worker_mem <- max(1L, floor(terra_mem_gb * 0.8 / n_workers))
cat(sprintf("Per-worker terra memmax: %d GB\n", per_worker_mem))

results <- parallel::mclapply(work_units, function(wu) {
  library(terra)
  library(minpack.lm)
  terraOptions(memmax = per_worker_mem)
  calc_cr_tile(
    tile_id      = wu$tile_id,
    input_dir    = input_dir,
    output_dir   = output_dir,
    ages         = ages,
    n_bands      = n_bands,
    log_file     = log_file,
    app_cores    = 1L,
    test_crop_n  = if (is.null(wu$crop_ext)) test_crop_n else 0L,
    crop_ext     = wu$crop_ext,
    strip_suffix = wu$strip_suffix
  )
}, mc.cores = n_workers)

# Check for errors
errors <- sapply(results, inherits, "try-error")
if (any(errors)) {
  warning(sprintf("%d of %d work units failed", sum(errors), length(results)))
}

toc()
gc()
```

---

## 7. Smoke Test Diagnostics

```{r diagnostics, fig.width=10, fig.height=8}
if (!test_mode) {
  message("Diagnostics only run in test_mode. Set test_mode <- TRUE to use.")
} else {
  library(ggplot2)

  # ── Reload the fitted parameter rasters ──────────────────────────────────────
  tile_id  <- g_ids[1]
  a_path   <- file.path(output_dir, "A",           sprintf("A_%s.tif",           tile_id))
  k_path   <- file.path(output_dir, "K",           sprintf("K_%s.tif",           tile_id))
  b_path   <- file.path(output_dir, "B",           sprintf("B_%s.tif",           tile_id))
  cnv_path <- file.path(output_dir, "convergence", sprintf("convergence_%s.tif", tile_id))

  A_rast   <- rast(a_path)
  K_rast   <- rast(k_path)
  B_rast   <- rast(b_path)
  cnv_rast <- rast(cnv_path)

  A_vals   <- values(A_rast,   mat = FALSE)
  K_vals   <- values(K_rast,   mat = FALSE)
  B_vals   <- values(B_rast,   mat = FALSE)
  cnv_vals <- values(cnv_rast, mat = FALSE)

  # ── Convergence summary ───────────────────────────────────────────────────────
  cnv_tbl <- table(cnv_vals, useNA = "ifany")
  cat("Convergence codes:\n")
  cat("  1 = first iteration | 2 = second | 3 = third | 4 = failed | NA = skipped\n")
  print(cnv_tbl)
  n_valid  <- sum(!is.na(A_vals) & A_vals != 0)
  n_failed <- sum(cnv_vals == 4, na.rm = TRUE)
  cat(sprintf("\nValid fits: %d / %d pixels (%.1f%% failure rate)\n",
              n_valid, length(A_vals),
              100 * n_failed / max(n_valid, 1)))

  # ── Sample pixels with successful fits for curve plots ───────────────────────
  valid_idx <- which(!is.na(A_vals) & A_vals > 0 & cnv_vals %in% 1:3)
  if (length(valid_idx) < 6) {
    message("Fewer than 6 successfully fitted pixels — adjust test_crop_n or tile.")
  } else {
    set.seed(42)
    sample_idx <- sample(valid_idx, 6)

    # Reload the stacked input for the same cropped tile
    age_files  <- file.path(input_dir, sprintf("age_%03d_%s.tif", ages, tile_id))
    agc_stack  <- rast(age_files)
    e_out      <- ext(A_rast)  # output extent = cropped extent
    agc_crop   <- crop(agc_stack, e_out)
    agc_mat    <- values(agc_crop, mat = TRUE)  # n_pixels × 400

    age_v <- rep(ages, each = n_bands)

    # Build a long data frame of observed + fitted values for each sample pixel
    plot_df <- lapply(seq_along(sample_idx), function(i) {
      pix  <- sample_idx[i]
      obs  <- data.frame(
        pixel = paste0("pixel ", i),
        age   = age_v,
        agc   = agc_mat[pix, ],
        type  = "observed"
      )
      t_seq   <- seq(5, 100, length.out = 200)
      fitted  <- data.frame(
        pixel = paste0("pixel ", i),
        age   = t_seq,
        agc   = chapman_richards(t_seq, A_vals[pix], K_vals[pix], B_vals[pix]),
        type  = "fitted"
      )
      rbind(obs, fitted)
    }) |> do.call(what = rbind)

    print(
      ggplot(plot_df, aes(x = age, y = agc, color = type)) +
        geom_point(data = subset(plot_df, type == "observed"),
                   alpha = 0.25, size = 0.8) +
        geom_line(data  = subset(plot_df, type == "fitted"),
                  linewidth = 1) +
        facet_wrap(~pixel, ncol = 3, scales = "free_y") +
        scale_color_manual(values = c(observed = "steelblue", fitted = "firebrick")) +
        labs(
          title = sprintf("CR curve fits — %s tile (test crop %d×%d px)",
                          scenario, test_crop_n, test_crop_n),
          x     = "Stand age (years)",
          y     = "AGC (Mg C / ha)",
          color = NULL
        ) +
        theme_minimal(base_size = 11) +
        theme(legend.position = "bottom")
    )
  }
}
```

---

## 8. (Optional) Mosaic Output Tiles  

After all tiles for a scenario are complete, mosaic the per-tile GeoTIFFs into
a single global raster for each parameter. Skip this section if you intend to
mosaic in a separate step or in GEE.

```{r mosaic, eval=FALSE}
mosaic_dir <- file.path(output_dir, "mosaic")
dir.create(mosaic_dir, recursive = TRUE, showWarnings = FALSE)

for (param in param_names) {
  tile_files <- list.files(file.path(output_dir, param),
                            pattern = "\\.tif$", full.names = TRUE)
  if (length(tile_files) == 0) {
    message(sprintf("No output tiles found for parameter '%s' — skipping.", param))
    next
  }

  tile_rasts <- lapply(tile_files, rast)
  mosaic_rast <- do.call(mosaic, c(tile_rasts, list(fun = "mean")))

  out_mosaic <- file.path(mosaic_dir,
                           sprintf("%s_%s_global.tif", param, scenario))
  writeRaster(mosaic_rast, out_mosaic,
              gdal = "COMPRESS=DEFLATE", overwrite = TRUE)
  message(sprintf("Wrote: %s", out_mosaic))
}
```

---

## 9. Run Summary

```{r summary}
# Count completed output tiles per parameter
summary_tbl <- data.frame(
  Parameter = param_names,
  Tiles_written = sapply(param_names, function(p) {
    length(list.files(file.path(output_dir, p), pattern = "\\.tif$"))
  })
)

knitr::kable(summary_tbl,
             caption = sprintf("Output tile counts — scenario: %s", scenario))
```

```{r log_tail}
# Show tail of progress log
log_lines <- readLines(log_file)
tail_n <- min(30, length(log_lines))
cat(tail(log_lines, tail_n), sep = "\n")
```

---

---

## 9. Sync Outputs to S3

Only runs when `save_to_s3 <- TRUE` in Section 2. Run this after all tiles
for the current scenario are complete. On SageMaker the instance IAM role
provides credentials automatically.

```{r sync-to-s3}
if (!save_to_s3) {
  message("save_to_s3 is FALSE — skipping S3 sync.")
} else {
  cat(sprintf("Syncing %s outputs to S3...\n", scenario))
  cat(sprintf("  From: %s\n", output_dir))
  cat(sprintf("  To:   %s\n", s3_output_dir))

  # rclone is used instead of aws CLI (not available in SageMaker Posit shell)
  ret <- system2(
    "rclone",
    c("copy", output_dir, s3_output_dir, "--transfers", "20", "--progress"),
    stdout = TRUE, stderr = TRUE
  )

  if (length(ret) > 0) cat(ret, sep = "\n")
  cat(sprintf("\nDone — %s outputs synced to S3.\n", scenario))
}
```

---

*To run the second scenario, change `scenario <- "hist"` in Section 2 and
re-run the notebook (Knit, or run chunks interactively).*
