---
title: "Chapman-Richards Curve Parameter Estimation"
subtitle: "Pixel-level AGC growth curve fitting for natural forest regeneration"
author: "Nathaniel Robinson / adapted by Serge Wiltshire"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: show
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
rm(list = ls())
```

---

## Overview

This notebook fits a **Chapman-Richards (CR)** growth curve to pixel-level above-ground
carbon (AGC) estimates from random forest outputs. For each pixel, the model is:

$$
\text{AGC}(t) = A \cdot \left(1 - B \cdot e^{-K \cdot t}\right)^{\frac{1}{1-m}}, \quad m = \tfrac{2}{3}
$$

which simplifies to:

$$
\text{AGC}(t) = A \cdot \left(1 - B \cdot e^{-K \cdot t}\right)^{3}
$$

**Parameters estimated per pixel:**

| Parameter | Meaning |
|-----------|---------|
| A | Asymptote — maximum potential AGC |
| K | Growth rate |
| B | Shape parameter |

**Inputs** are GEE-exported tiles (one per 62°×62° region) at ~1 km resolution:

- `data/inputs/{scenario}/age_{AGE}_{TILE_ID}.tif` — 20-band rasters (random forest
  realizations `sd_001`–`sd_020`) at 20 age steps (5, 10, …, 100 years)
- **future** and **hist** scenarios are run separately by changing `scenario` below.

**Outputs** are written to `data/outputs/{scenario}/` in subdirectories `A/`, `K/`, `B/`,
`A_error/`, `K_error/`, `B_error/`, `convergence/`.

> **Computational note:** Each tile is 7 424 × 7 424 pixels with 400 bands stacked
> (20 ages × 20 realizations). Tiles are dispatched in parallel (one `mclapply`
> worker per tile). Fitting uses `terra::app()` so rasters are read and written
> in memory-efficient chunks. On SageMaker, run via the wrapper script from the
> terminal (`nohup Rscript cr_calc/cr_chapman_richards_wrapper.R`) to avoid
> mclapply fork crashes inside RStudio Server.

---

## 1. Package Setup

```{r packages}
required_packages <- c(
  "terra",
  "tidyverse",
  "minpack.lm",
  "parallel"
)

new_pkgs <- required_packages[
  !(required_packages %in% installed.packages()[, "Package"])
]
if (length(new_pkgs) > 0) {
  install.packages(new_pkgs, dep = TRUE)
}

invisible(lapply(required_packages, library, character.only = TRUE))
```

---

## 2. Configuration

Adjust these parameters before each run.

```{r config}
# ── Smoke test toggle ─────────────────────────────────────────────────────────
# TRUE  → run on one tile cropped to test_crop_n × test_crop_n pixels.
#         Completes in minutes on a laptop. Use this to verify the fitting
#         pipeline and inspect example curves before a full production run.
# FALSE → full production run over all tiles.
test_mode    <- FALSE
test_crop_n  <- 150   # pixels per side when test_mode = TRUE

# ── Scenario ──────────────────────────────────────────────────────────────────
# Scenarios to process. Use c("hist", "future") to run both in parallel
# (doubles worker count to 36 but halves total wall time).
scenarios <- c("hist", "future")

# ── Paths ─────────────────────────────────────────────────────────────────────
# Notebook-relative root of the project (two levels up from cr_calc/).
# Detect project root: use rstudioapi if available (interactive RStudio),
# otherwise fall back to the working directory (Rscript / command line).
if (requireNamespace("rstudioapi", quietly = TRUE) && rstudioapi::isAvailable()) {
  project_root <- normalizePath(file.path(
    dirname(rstudioapi::getActiveDocumentContext()$path), ".."))
} else {
  project_root <- getwd()
}

# Paths are built per-scenario in the work_units chunk (Section 4).

# ── Age sequence ──────────────────────────────────────────────────────────────
# Ages (years) matching the input file naming convention (age_005, age_010, …)
ages    <- seq(5, 100, by = 5)   # 20 time steps
n_bands <- 20                    # realizations per age file (sd_001 – sd_020)

# Predictor vector for NLS: each age is repeated once per realization band,
# producing a length-400 vector aligned to the stacked 400-band raster.
age_vector <- rep(ages, each = n_bands)

# ── Parallelisation ────────────────────────────────────────────────────────────
# One mclapply worker per work unit (= one tile). Each worker processes a
# whole tile via app() which reads lazily in chunks. Each worker reads its
# own set of 20 input files — no I/O contention between workers.
n_tiles <- 18L
n_cores <- max(1L, detectCores() - 1L)

cat(sprintf("Cores: %d | Tiles per scenario: %d\n", n_cores, n_tiles))

# ── Memory ────────────────────────────────────────────────────────────────────
# Total terra memory budget (GB) shared across all workers.
# ml.r5.12xlarge = 384 GB: use ~250 GB, leaving headroom for R + OS.
# Each mclapply worker sets its own memmax = floor(terra_mem_gb / n_workers).
terra_mem_gb <- 250
terraOptions(memmax = terra_mem_gb)

# ── S3 output ─────────────────────────────────────────────────────────────────
# TRUE  → after all tiles complete, sync local outputs to S3 (Section 9).
#          Requires rclone configured (see setup/fetch_input_data.Rmd).
# FALSE → outputs stay local only (laptop / smoke test).
save_to_s3 <- TRUE

s3_output_base <- "s3://sagemaker-gst-stage.sharing/serge-wiltshire/nat_for_regen_c_accumulation_data"
```

---

## 3. Function Definitions

### 3.1 Chapman-Richards model

```{r fn_cr}
#' Chapman-Richards growth model (m fixed at 2/3)
#'
#' @param t  Numeric vector of ages (years).
#' @param A  Asymptote parameter.
#' @param K  Growth rate parameter.
#' @param B  Shape parameter.
#' @return   Numeric vector of predicted AGC values.
chapman_richards <- function(t, A, K, B) {
  A * (1 - B * exp(-K * t))^3
}
```

### 3.2 Single NLS attempt

```{r fn_fit_nls}
#' Attempt one NLS fit of the CR curve.
#'
#' @param pix_y   Numeric response vector (AGC values for one pixel,
#'                length = n_ages * n_bands = 400).
#' @param age_v   Numeric predictor vector of the same length as pix_y.
#' @param a_start Starting value for A (typically max_pot for this pixel).
#' @param k_start Starting value for K.
#' @param b_start Starting value for B.
#' @return Named numeric vector (A, K, B, A_err, K_err, B_err) on success,
#'         or NULL on failure.
attempt_nls <- function(pix_y, age_v,
                        a_start,
                        k_start = 0.05,
                        b_start = 0.75) {
  lower <- c(a = a_start * 0.9, k = 0.01, b = 0.2)
  upper <- c(a = a_start * 1.1, k = 0.10, b = 1.0)

  fit <- try(
    minpack.lm::nlsLM(
      pix_y ~ a * (1 - b * exp(-k * age_v))^3,
      start   = list(a = a_start, k = k_start, b = b_start),
      lower   = lower,
      upper   = upper,
      control = minpack.lm::nls.lm.control(maxiter = 500)
    ),
    silent = TRUE
  )

  if (inherits(fit, "try-error")) return(NULL)

  cf <- coef(fit)
  se <- tryCatch(
    summary(fit)$coefficients[, "Std. Error"],
    error = function(e) rep(NA_real_, 3)
  )

  c(A     = unname(cf["a"]),
    K     = unname(cf["k"]),
    B     = unname(cf["b"]),
    A_err = se[1],
    K_err = se[2],
    B_err = se[3])
}
```

### 3.3 Multi-iteration pixel fitting

```{r fn_fit_pixel}
#' Fit CR curve for a single pixel with up to 3 NLS iterations.
#'
#' Iteration strategy:
#'   1. Start from max_pot + default k/b.
#'   2. If iteration 1 did not fully converge (stopCode ≠ 0),
#'      re-start from iteration 1 coefficients.
#'   3. If iteration 2 still did not converge, re-start from iteration 2
#'      coefficients.
#'   4. If all three fail, return NA with convergence code 4.
#'
#' Convergence codes: 1 = converged on 1st try, 2 = 2nd, 3 = 3rd, 4 = failed.
#'
#' @param pix_y   Numeric response vector (length = n_ages * n_bands).
#' @param age_v   Numeric predictor vector (same length as pix_y).
#' @param max_pot Maximum potential AGC for this pixel (used to anchor A).
#' @return Named numeric vector of length 7:
#'         A, K, B, A_err, K_err, B_err, convergence.
fit_cr_pixel <- function(pix_y, age_v, max_pot) {
  na_result  <- c(A = NA_real_, K = NA_real_, B = NA_real_,
                  A_err = NA_real_, K_err = NA_real_, B_err = NA_real_,
                  convergence = NA_real_)
  zero_result <- c(A = 0, K = 0, B = 0,
                   A_err = 0, K_err = 0, B_err = 0,
                   convergence = 0)

  if (all(is.na(pix_y))) return(na_result)

  # Filter out NA bands and their corresponding age entries
  valid <- !is.na(pix_y)
  pix_y <- pix_y[valid]
  age_v <- age_v[valid]

  # Need enough data: at least 5 distinct age steps represented
  if (length(unique(age_v)) < 5) return(na_result)

  max_pot <- max(pix_y)
  if (max_pot <= 0) return(zero_result)

  # Helper: check whether nls converged (stopCode == 0 for "port" algorithm)
  converged <- function(fit) {
    tryCatch(
      isTRUE(fit$convInfo$stopCode == 0),
      error = function(e) TRUE  # assume ok if info unavailable
    )
  }

  # ── Iteration 1 ─────────────────────────────────────────────────────────────
  r1 <- attempt_nls(pix_y, age_v, a_start = max_pot)
  if (is.null(r1)) return(c(na_result[1:6], convergence = 4))

  # Retrieve the fit object again briefly just to check stopCode
  # (attempt_nls already ran it; we skip re-running here and use warnOnly
  #  heuristic: if A/K/B are within bounds, treat as converged enough)
  if (r1["A"] >= max_pot * 0.9 && r1["A"] <= max_pot * 1.1) {
    return(c(r1, convergence = 1))
  }

  # ── Iteration 2: re-start from iteration 1 result ───────────────────────────
  r2 <- attempt_nls(pix_y, age_v,
                    a_start = r1["A"],
                    k_start = r1["K"],
                    b_start = r1["B"])
  if (is.null(r2)) return(c(r1, convergence = 4))

  if (r2["A"] >= r1["A"] * 0.9 && r2["A"] <= r1["A"] * 1.1) {
    return(c(r2, convergence = 2))
  }

  # ── Iteration 3: re-start from iteration 2 result ───────────────────────────
  r3 <- attempt_nls(pix_y, age_v,
                    a_start = r2["A"],
                    k_start = r2["K"],
                    b_start = r2["B"])
  if (is.null(r3)) return(c(r2, convergence = 4))

  c(r3, convergence = 3)
}
```

### 3.4 Tile-level CR calculation

```{r fn_calc_tile}
#' Fit CR curves for all pixels in a single tile.
#'
#' Stacks the 400-band raster (20 ages × 20 realizations) for the given tile,
#' then applies pixel-wise NLS fitting via terra::app() — raster chunks are
#' read sequentially so the full tile never lives in RAM at once.
#' Outputs are written as compressed single-band GeoTIFFs.
#'
#' @param tile_id   Character tile identifier extracted from filenames
#'                  (e.g. "17817146248550000000000-0000000000").
#' @param input_dir Full path to the scenario input directory.
#' @param output_dir Full path to the scenario output directory.
#' @param ages      Numeric vector of age steps (default seq(5, 100, by=5)).
#' @param n_bands   Number of realization bands per age file (default 20).
#' @param log_file  Path to log file for progress messages.
#' @param app_cores   Integer number of cores to pass to terra::app().
#'                    Use n_cores in "pixel" strategy; 1L in "tile" strategy.
#' @param test_crop_n  Integer. If > 0, auto-locate valid data and crop to a
#'                     test_crop_n x test_crop_n pixel window. Smoke test only.
#' @param crop_ext     Numeric vector c(xmin, xmax, ymin, ymax) or NULL.
#'                     If provided, crop the stacked raster to this extent
#'                     before fitting. Takes precedence over test_crop_n.
#' @param strip_suffix Character suffix for output filenames (default "").
#' @return Invisibly returns the tile_id on success.
calc_cr_tile <- function(tile_id,
                         input_dir,
                         output_dir,
                         ages         = seq(5, 100, by = 5),
                         n_bands      = 20,
                         log_file     = NULL,
                         app_cores    = 1L,
                         test_crop_n  = 0L,
                         crop_ext     = NULL,
                         strip_suffix = "") {

  label <- paste0(tile_id, strip_suffix)
  pid   <- Sys.getpid()

  # Helper: write to log file (append, atomic-ish for concurrent workers)
  log_msg <- function(step, detail = "") {
    msg <- sprintf("[%s | pid=%d] %s %s — %s\n",
                   label, pid, step, detail, format(Sys.time(), "%H:%M:%S"))
    if (!is.null(log_file)) cat(msg, file = log_file, append = TRUE)
  }

  age_v <- rep(ages, each = n_bands)

  # ── Build file list ──────────────────────────────────────────────────────────
  log_msg("STEP 1/6", "building file list")
  age_files <- file.path(input_dir, sprintf("age_%03d_%s.tif", ages, tile_id))

  missing_files <- age_files[!file.exists(age_files)]
  if (length(missing_files) > 0) {
    log_msg("SKIPPED", sprintf("%d missing files", length(missing_files)))
    return(invisible(NULL))
  }

  # ── Stack raster ─────────────────────────────────────────────────────────────
  log_msg("STEP 2/6", "stacking 400-band raster")
  agc_stack <- rast(age_files)

  # ── Crop ─────────────────────────────────────────────────────────────────────
  if (!is.null(crop_ext)) {
    log_msg("STEP 3/6", sprintf("cropping to extent [%.2f, %.2f, %.2f, %.2f]",
                                crop_ext[1], crop_ext[2], crop_ext[3], crop_ext[4]))
    agc_stack <- crop(agc_stack, ext(crop_ext))
  } else if (test_crop_n > 0) {
    log_msg("STEP 3/6", sprintf("auto-locating %dx%d test crop", test_crop_n, test_crop_n))
    probe   <- rast(age_files[length(age_files)])[[1]]
    agg_fct <- max(1L, floor(min(nrow(probe), ncol(probe)) / 75L))
    probe_agg <- aggregate(probe, fact = agg_fct, fun = "max", na.rm = TRUE)
    agg_vals  <- values(probe_agg, mat = FALSE)
    best_cell <- which.max(replace(agg_vals, is.na(agg_vals), 0))
    if (length(best_cell) == 0 || max(agg_vals, na.rm = TRUE) == 0) {
      e  <- ext(agc_stack)
      rx <- res(agc_stack)[1]; ry <- res(agc_stack)[2]
      crop_ext <- ext(e[1], e[1] + test_crop_n*rx, e[4] - test_crop_n*ry, e[4])
    } else {
      best_row <- rowFromCell(probe_agg, best_cell)
      best_col <- colFromCell(probe_agg, best_cell)
      mid_row  <- (best_row - 0.5) * agg_fct
      mid_col  <- (best_col - 0.5) * agg_fct
      e  <- ext(agc_stack)
      rx <- res(agc_stack)[1]; ry <- res(agc_stack)[2]
      half <- test_crop_n / 2
      x_mid <- e[1] + (mid_col - 0.5) * rx
      y_mid <- e[4] - (mid_row - 0.5) * ry
      crop_ext <- ext(x_mid - half*rx, x_mid + half*rx,
                      y_mid - half*ry, y_mid + half*ry)
    }
    agc_stack <- crop(agc_stack, crop_ext)
  } else {
    log_msg("STEP 3/6", "no crop (full tile)")
  }

  log_msg("STEP 3/6", sprintf("raster size: %d x %d px, %d bands",
                               nrow(agc_stack), ncol(agc_stack), nlyr(agc_stack)))

  # ── Build self-contained pixel fitting function ──────────────────────────────
  log_msg("STEP 4/6", "building pixel fitting closure")
  cr_app_func <- local({
    av <- age_v
    .attempt_nls <- function(pix_y, age_v, a_start, k_start = 0.05, b_start = 0.75) {
      lower <- c(a = a_start * 0.9, k = 0.01, b = 0.2)
      upper <- c(a = a_start * 1.1, k = 0.10, b = 1.0)
      fit <- try(minpack.lm::nlsLM(
        pix_y ~ a * (1 - b * exp(-k * age_v))^3,
        start   = list(a = a_start, k = k_start, b = b_start),
        lower   = lower, upper = upper,
        control = minpack.lm::nls.lm.control(maxiter = 500)
      ), silent = TRUE)
      if (inherits(fit, "try-error")) return(NULL)
      cf <- coef(fit)
      se <- tryCatch(summary(fit)$coefficients[, "Std. Error"],
                     error = function(e) rep(NA_real_, 3))
      c(A = unname(cf["a"]), K = unname(cf["k"]), B = unname(cf["b"]),
        A_err = se[1], K_err = se[2], B_err = se[3])
    }
    .fit_cr_pixel <- function(pix_y, age_v, max_pot) {
      na_r <- c(A=NA_real_,K=NA_real_,B=NA_real_,A_err=NA_real_,K_err=NA_real_,B_err=NA_real_,convergence=NA_real_)
      if (all(is.na(pix_y))) return(na_r)

      # Filter NA bands and corresponding age entries
      valid <- !is.na(pix_y)
      pix_y <- pix_y[valid]
      age_v <- age_v[valid]
      if (length(unique(age_v)) < 5) return(na_r)

      max_pot <- max(pix_y)
      if (max_pot <= 0) return(c(A=0,K=0,B=0,A_err=0,K_err=0,B_err=0,convergence=0))
      r1 <- .attempt_nls(pix_y, age_v, a_start = max_pot)
      if (is.null(r1)) return(c(na_r[1:6], convergence = 4))
      if (r1["A"] >= max_pot*0.9 & r1["A"] <= max_pot*1.1) return(c(r1, convergence=1))
      r2 <- .attempt_nls(pix_y, age_v, a_start=r1["A"], k_start=r1["K"], b_start=r1["B"])
      if (is.null(r2)) return(c(r1, convergence=4))
      if (r2["A"] >= r1["A"]*0.9 & r2["A"] <= r1["A"]*1.1) return(c(r2, convergence=2))
      r3 <- .attempt_nls(pix_y, age_v, a_start=r2["A"], k_start=r2["K"], b_start=r2["B"])
      if (is.null(r3)) return(c(r2, convergence=4))
      c(r3, convergence=3)
    }
    function(x) {
      .fit_cr_pixel(pix_y = x, age_v = av, max_pot = max(x, na.rm = TRUE))
    }
  })

  # ── Run pixel-wise fitting ──────────────────────────────────────────────────
  log_msg("STEP 5/6", sprintf("starting app() — %d total pixels",
                               nrow(agc_stack) * ncol(agc_stack)))
  result <- app(agc_stack, fun = cr_app_func, cores = app_cores)
  names(result) <- c("A", "K", "B", "A_error", "K_error", "B_error", "convergence")
  log_msg("STEP 5/6", "app() complete")

  # ── Write outputs ───────────────────────────────────────────────────────────
  log_msg("STEP 6/6", "writing GeoTIFFs")
  param_dirs <- file.path(output_dir, names(result))
  invisible(lapply(param_dirs, dir.create, recursive = TRUE, showWarnings = FALSE))

  for (param in names(result)) {
    out_path <- file.path(output_dir, param,
                          sprintf("%s_%s%s.tif", param, tile_id, strip_suffix))
    writeRaster(result[[param]], out_path, gdal = "COMPRESS=DEFLATE", overwrite = TRUE)
  }

  log_msg("DONE", "all outputs written")
  invisible(tile_id)
}
```

---

## 4. File Discovery

```{r work_units}
if (test_mode) {

  # ── Smoke test: one tile from the first scenario, auto-locate crop ──────────
  scn       <- scenarios[1]
  input_dir <- file.path(project_root, "data", "inputs",  scn)
  all_files <- list.files(input_dir, pattern = "\\.tif$")
  all_tile_ids <- unique(sub("^age_\\d+_(.+)\\.tif$", "\\1", all_files))

  work_units <- list(list(
    tile_id      = all_tile_ids[1],
    crop_ext     = NULL,
    strip_suffix = "",
    input_dir    = input_dir,
    output_dir   = file.path(project_root, "data", "outputs", scn),
    log_file     = file.path(project_root, "data", "outputs", scn,
                             sprintf("progress_%s.log", scn)),
    scenario     = scn
  ))
  cat(sprintf("[TEST MODE] 1 work unit (%s, 1 tile × %d×%d px crop)\n",
              scn, test_crop_n, test_crop_n))

} else {

  # ── Production: one work unit per tile × scenario ───────────────────────────
  work_units <- unlist(lapply(scenarios, function(scn) {
    input_dir  <- file.path(project_root, "data", "inputs",  scn)
    output_dir <- file.path(project_root, "data", "outputs", scn)
    log_file   <- file.path(output_dir, sprintf("progress_%s.log", scn))

    all_files    <- list.files(input_dir, pattern = "\\.tif$")
    tile_ids     <- unique(sub("^age_\\d+_(.+)\\.tif$", "\\1", all_files))

    cat(sprintf("Scenario: %s | Tiles: %d | Input: %s\n",
                scn, length(tile_ids), input_dir))

    # Skip completed tiles
    completed_dir <- file.path(output_dir, "A")
    completed_files <- if (dir.exists(completed_dir))
      list.files(completed_dir, pattern = "\\.tif$") else character(0)

    lapply(tile_ids, function(tid) {
      if (sprintf("A_%s.tif", tid) %in% completed_files) return(NULL)
      list(tile_id = tid, crop_ext = NULL, strip_suffix = "",
           input_dir = input_dir, output_dir = output_dir,
           log_file = log_file, scenario = scn)
    })
  }), recursive = FALSE)

  # Remove NULLs (completed tiles)
  work_units <- Filter(Negate(is.null), work_units)

  cat(sprintf("\nTotal work units: %d across %d scenario(s)\n",
              length(work_units), length(scenarios)))
}
```

---

## 5. Output Directory Setup

```{r output_dirs}
param_names <- c("A", "K", "B", "A_error", "K_error", "B_error", "convergence")

# Create output dirs and log files for each scenario
for (scn in scenarios) {
  out_dir <- file.path(project_root, "data", "outputs", scn)
  invisible(lapply(file.path(out_dir, param_names),
                   dir.create, recursive = TRUE, showWarnings = FALSE))
  lf <- file.path(out_dir, sprintf("progress_%s.log", scn))
  file.create(lf, showWarnings = FALSE)
  cat(sprintf("Run started: %s\nScenario: %s\n",
              format(Sys.time()), scn), file = lf, append = TRUE)
  cat(sprintf("Output directories ready: %s\n", out_dir))
}
```

---

## 6. CR Parameter Fitting

Parallel via `mclapply` (fork-based on Linux). One worker per work unit,
each reading its own files with zero I/O contention.

On SageMaker, run via the wrapper script from the terminal — **not** as an
RStudio background job (mclapply fork crashes inside RStudio Server).

Monitor both scenario logs simultaneously:
`tail -f data/outputs/*/progress_*.log`

```{r fit_parallel}
n_workers <- min(n_cores, length(work_units))
per_worker_mem <- max(1L, floor(terra_mem_gb / n_workers))

cat(sprintf("=== CR fitting ===\n"))
cat(sprintf("  Scenarios:  %s\n", paste(scenarios, collapse = ", ")))
cat(sprintf("  Work units: %d\n", length(work_units)))
cat(sprintf("  Workers:    %d (of %d cores)\n", n_workers, parallel::detectCores()))
cat(sprintf("  Per-worker terra memmax: %d GB\n", per_worker_mem))
cat(sprintf("  Started: %s\n", format(Sys.time())))

results <- parallel::mclapply(work_units, function(wu) {
  library(terra)
  library(minpack.lm)
  terraOptions(memmax = per_worker_mem)
  tryCatch(
    calc_cr_tile(
      tile_id      = wu$tile_id,
      input_dir    = wu$input_dir,
      output_dir   = wu$output_dir,
      ages         = ages,
      n_bands      = n_bands,
      log_file     = wu$log_file,
      app_cores    = 1L,
      test_crop_n  = if (is.null(wu$crop_ext)) test_crop_n else 0L,
      crop_ext     = wu$crop_ext,
      strip_suffix = wu$strip_suffix
    ),
    error = function(e) {
      msg <- sprintf("[%s/%s | pid=%d] FATAL: %s\n",
                     wu$scenario, wu$tile_id, Sys.getpid(), conditionMessage(e))
      cat(msg, file = wu$log_file, append = TRUE)
      paste("ERROR:", conditionMessage(e))
    }
  )
}, mc.cores = n_workers)

# ── Report ────────────────────────────────────────────────────────────────
is_err <- vapply(results, function(x) {
  is.character(x) && startsWith(x, "ERROR:")
}, logical(1))
cat(sprintf("\n=== Done | %d work units | %d errors | %s ===\n",
            length(results), sum(is_err), format(Sys.time())))

if (any(is_err)) {
  cat("\nErrors:\n")
  invisible(lapply(results[is_err], function(e) cat("  ", e, "\n")))
}
```

---

## 7. Smoke Test Diagnostics

```{r diagnostics, fig.width=10, fig.height=8}
if (!test_mode) {
  message("Diagnostics only run in test_mode. Set test_mode <- TRUE to use.")
} else {
  library(ggplot2)

  # ── Reload the fitted parameter rasters ──────────────────────────────────────
  wu         <- work_units[[1]]
  tile_id    <- wu$tile_id
  suffix     <- wu$strip_suffix
  output_dir <- wu$output_dir
  input_dir  <- wu$input_dir
  a_path   <- file.path(output_dir, "A",           sprintf("A_%s%s.tif",           tile_id, suffix))
  k_path   <- file.path(output_dir, "K",           sprintf("K_%s%s.tif",           tile_id, suffix))
  b_path   <- file.path(output_dir, "B",           sprintf("B_%s%s.tif",           tile_id, suffix))
  cnv_path <- file.path(output_dir, "convergence", sprintf("convergence_%s%s.tif", tile_id, suffix))

  A_rast   <- rast(a_path)
  K_rast   <- rast(k_path)
  B_rast   <- rast(b_path)
  cnv_rast <- rast(cnv_path)

  A_vals   <- values(A_rast,   mat = FALSE)
  K_vals   <- values(K_rast,   mat = FALSE)
  B_vals   <- values(B_rast,   mat = FALSE)
  cnv_vals <- values(cnv_rast, mat = FALSE)

  # ── Convergence summary ───────────────────────────────────────────────────────
  cnv_tbl <- table(cnv_vals, useNA = "ifany")
  cat("Convergence codes:\n")
  cat("  1 = first iteration | 2 = second | 3 = third | 4 = failed | NA = skipped\n")
  print(cnv_tbl)
  n_valid  <- sum(!is.na(A_vals) & A_vals != 0)
  n_failed <- sum(cnv_vals == 4, na.rm = TRUE)
  cat(sprintf("\nValid fits: %d / %d pixels (%.1f%% failure rate)\n",
              n_valid, length(A_vals),
              100 * n_failed / max(n_valid, 1)))

  # ── Sample pixels with successful fits for curve plots ───────────────────────
  valid_idx <- which(!is.na(A_vals) & A_vals > 0 & cnv_vals %in% 1:3)
  if (length(valid_idx) < 6) {
    message("Fewer than 6 successfully fitted pixels — adjust test_crop_n or tile.")
  } else {
    set.seed(42)
    sample_idx <- sample(valid_idx, 6)

    # Reload the stacked input for the same cropped tile
    age_files  <- file.path(input_dir, sprintf("age_%03d_%s.tif", ages, tile_id))
    agc_stack  <- rast(age_files)
    e_out      <- ext(A_rast)  # output extent = cropped extent
    agc_crop   <- crop(agc_stack, e_out)
    agc_mat    <- values(agc_crop, mat = TRUE)  # n_pixels × 400

    age_v <- rep(ages, each = n_bands)

    # Build a long data frame of observed + fitted values for each sample pixel
    plot_df <- lapply(seq_along(sample_idx), function(i) {
      pix  <- sample_idx[i]
      obs  <- data.frame(
        pixel = paste0("pixel ", i),
        age   = age_v,
        agc   = agc_mat[pix, ],
        type  = "observed"
      )
      t_seq   <- seq(5, 100, length.out = 200)
      fitted  <- data.frame(
        pixel = paste0("pixel ", i),
        age   = t_seq,
        agc   = chapman_richards(t_seq, A_vals[pix], K_vals[pix], B_vals[pix]),
        type  = "fitted"
      )
      rbind(obs, fitted)
    }) |> do.call(what = rbind)

    print(
      ggplot(plot_df, aes(x = age, y = agc, color = type)) +
        geom_point(data = subset(plot_df, type == "observed"),
                   alpha = 0.25, size = 0.8) +
        geom_line(data  = subset(plot_df, type == "fitted"),
                  linewidth = 1) +
        facet_wrap(~pixel, ncol = 3, scales = "free_y") +
        scale_color_manual(values = c(observed = "steelblue", fitted = "firebrick")) +
        labs(
          title = sprintf("CR curve fits — %s tile (test crop %d×%d px)",
                          scenario, test_crop_n, test_crop_n),
          x     = "Stand age (years)",
          y     = "AGC (Mg C / ha)",
          color = NULL
        ) +
        theme_minimal(base_size = 11) +
        theme(legend.position = "bottom")
    )
  }
}
```

---

## 8. (Optional) Mosaic Output Tiles  

After all tiles for a scenario are complete, mosaic the per-tile GeoTIFFs into
a single global raster for each parameter. Skip this section if you intend to
mosaic in a separate step or in GEE.

```{r mosaic, eval=FALSE}
for (scn in scenarios) {
  out_dir    <- file.path(project_root, "data", "outputs", scn)
  mosaic_dir <- file.path(out_dir, "mosaic")
  dir.create(mosaic_dir, recursive = TRUE, showWarnings = FALSE)

  for (param in param_names) {
    tile_files <- list.files(file.path(out_dir, param),
                              pattern = "\\.tif$", full.names = TRUE)
    if (length(tile_files) == 0) next
    tile_rasts  <- lapply(tile_files, rast)
    mosaic_rast <- do.call(mosaic, c(tile_rasts, list(fun = "mean")))
    out_mosaic  <- file.path(mosaic_dir, sprintf("%s_%s_global.tif", param, scn))
    writeRaster(mosaic_rast, out_mosaic, gdal = "COMPRESS=DEFLATE", overwrite = TRUE)
    message(sprintf("Wrote: %s", out_mosaic))
  }
}
```

---

## 9. Run Summary

```{r summary}
for (scn in scenarios) {
  out_dir <- file.path(project_root, "data", "outputs", scn)
  counts  <- sapply(param_names, function(p)
    length(list.files(file.path(out_dir, p), pattern = "\\.tif$")))

  cat(sprintf("\n%s: %d / 18 tiles complete\n", toupper(scn), min(counts)))

  lf <- file.path(out_dir, sprintf("progress_%s.log", scn))
  if (file.exists(lf)) {
    log_lines <- readLines(lf)
    cat(tail(log_lines, 10), sep = "\n")
    cat("\n")
  }
}
```

---

---

## 9. Sync Outputs to S3

Only runs when `save_to_s3 <- TRUE` in Section 2. Run this after all tiles
for the current scenario are complete. On SageMaker the instance IAM role
provides credentials automatically.

```{r sync-to-s3}
if (!save_to_s3) {
  message("save_to_s3 is FALSE — skipping S3 sync.")
} else {
  for (scn in scenarios) {
    out_dir <- file.path(project_root, "data", "outputs", scn)
    s3_dest <- sprintf("%s/%s/outputs", s3_output_base, scn)
    cat(sprintf("Syncing %s: %s -> %s\n", scn, out_dir, s3_dest))
    system2("rclone", c("copy", out_dir, s3_dest, "--transfers", "20", "--progress"),
            stdout = TRUE, stderr = TRUE)
    cat(sprintf("Done: %s\n", scn))
  }
}
```

---

*Both scenarios run in a single launch when `scenarios <- c("hist", "future")`.*
