---
title: "Interpolate CR Parameters Across Future Years"
subtitle: "Linear interpolation of A, B, K between hist (2005) and future (2090)"
author: "Serge Wiltshire"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: show
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
rm(list = ls())
```

---

## Overview

This notebook linearly interpolates the Chapman-Richards parameters **A**, **B**, and
**K** between historical (2005) and future (2090) climate scenarios, producing
GeoTIFFs at 5-year intervals.

For each parameter $P \in \{A, B, K\}$ and each pixel:

$$
P(\text{year}) = P_{\text{hist}} + (P_{\text{future}} - P_{\text{hist}}) \times \frac{\text{year} - 2005}{85}
$$

where year $\in \{2005, 2010, 2015, \ldots, 2090\}$.

**Inputs:** single-band GeoTIFFs from `data/outputs/hist/{A,B,K}/` and
`data/outputs/future/{A,B,K}/` produced by `cr_chapman_richards.Rmd`.

**Outputs:** 18-band GeoTIFFs in `data/outputs/interpolated/{A,B,K}/`, one per tile,
with bands named `yr_2005`, `yr_2010`, `yr_2015`, …, `yr_2090`.

> **Computational note:** This is pure raster algebra (no NLS fitting) so each tile
> completes in seconds. Still runs via the wrapper script on SageMaker for mclapply
> fork safety.

---

## 1. Package Setup

```{r packages}
required_packages <- c("terra", "parallel")

new_pkgs <- required_packages[
  !(required_packages %in% installed.packages()[, "Package"])
]
if (length(new_pkgs) > 0) install.packages(new_pkgs, dep = TRUE)

invisible(lapply(required_packages, library, character.only = TRUE))
```

---

## 2. Configuration

```{r config}
# ── Smoke test toggle ─────────────────────────────────────────────────────────
# TRUE  → run on one tile cropped to test_crop_n × test_crop_n pixels.
# FALSE → full production run over all tiles.
test_mode   <- FALSE
test_crop_n <- 150L

# ── Paths ─────────────────────────────────────────────────────────────────────
if (requireNamespace("rstudioapi", quietly = TRUE) && rstudioapi::isAvailable()) {
  project_root <- normalizePath(file.path(
    dirname(rstudioapi::getActiveDocumentContext()$path), ".."))
} else {
  project_root <- getwd()
}

hist_dir   <- file.path(project_root, "data", "outputs", "hist")
future_dir <- file.path(project_root, "data", "outputs", "future")
output_dir <- file.path(project_root, "data", "outputs", "interpolated")

# ── Interpolation grid ────────────────────────────────────────────────────────
# 2005–2090 = 85 years; 5-year intervals divide evenly (18 steps)
years      <- seq(2005, 2090, by = 5)
weights    <- (years - 2005) / 85        # 0 = pure hist (2005), 1 = pure future (2090)
band_names <- sprintf("yr_%d", years)

# ── Parameters to interpolate ────────────────────────────────────────────────
params <- c("A", "B", "K")

# ── Parallelisation ──────────────────────────────────────────────────────────
n_cores <- max(1L, detectCores() - 1L)

# ── Memory ────────────────────────────────────────────────────────────────────
# ml.r5.12xlarge = 384 GB; allocate ~250 GB to terra.
terra_mem_gb <- 250
terraOptions(memmax = terra_mem_gb)

# ── S3 output ─────────────────────────────────────────────────────────────────
save_to_s3 <- TRUE
s3_output_base <- "s3://sagemaker-gst-stage.sharing/serge-wiltshire/nat_for_regen_c_accumulation_data"

cat(sprintf("Cores: %d\n", n_cores))
cat(sprintf("Years: %s\n", paste(years, collapse = ", ")))
```

---

## 3. Core Function

```{r fn_interpolate_tile}
#' Interpolate A, B, K for a single tile across decadal future years.
#'
#' Reads single-band hist and future parameter tiles, linearly interpolates
#' across the year grid, and writes one 10-band GeoTIFF per parameter.
#'
#' @param tile_id    Character tile identifier.
#' @param hist_dir   Path to hist output directory (contains A/, B/, K/).
#' @param future_dir Path to future output directory.
#' @param output_dir Path to interpolated output directory.
#' @param params     Character vector of parameter names (default c("A","B","K")).
#' @param weights    Numeric vector of interpolation weights (0 = hist, 1 = future).
#' @param band_names Character vector of band names for the output raster.
#' @param log_file   Path to progress log (NULL to skip logging).
#' @param test_crop_n Integer; if > 0, crop to a small window for testing.
#' @return Invisibly returns tile_id on success.
interpolate_tile <- function(tile_id,
                             hist_dir,
                             future_dir,
                             output_dir,
                             params     = c("A", "B", "K"),
                             weights    = (seq(2005, 2090, by = 5) - 2005) / 85,
                             band_names = sprintf("yr_%d", seq(2005, 2090, by = 5)),
                             log_file   = NULL,
                             test_crop_n = 0L) {

  pid <- Sys.getpid()

  log_msg <- function(step, detail = "") {
    msg <- sprintf("[%s | pid=%d] %s %s — %s\n",
                   tile_id, pid, step, detail, format(Sys.time(), "%H:%M:%S"))
    if (!is.null(log_file)) cat(msg, file = log_file, append = TRUE)
  }

  log_msg("START", sprintf("interpolating %d params × %d years",
                            length(params), length(weights)))

  for (param in params) {
    hist_path   <- file.path(hist_dir,   param, sprintf("%s_%s.tif", param, tile_id))
    future_path <- file.path(future_dir, param, sprintf("%s_%s.tif", param, tile_id))

    if (!file.exists(hist_path) || !file.exists(future_path)) {
      log_msg("SKIP", sprintf("%s — missing input", param))
      next
    }

    r_hist   <- rast(hist_path)
    r_future <- rast(future_path)

    # Optional test crop
    if (test_crop_n > 0) {
      e  <- ext(r_hist)
      rx <- res(r_hist)[1]; ry <- res(r_hist)[2]
      crop_ext <- ext(e[1], e[1] + test_crop_n * rx,
                      e[4] - test_crop_n * ry, e[4])
      r_hist   <- crop(r_hist,   crop_ext)
      r_future <- crop(r_future, crop_ext)
    }

    # Build 10-band stack via raster algebra
    r_diff  <- r_future - r_hist
    bands   <- lapply(weights, function(w) r_hist + r_diff * w)
    r_stack <- rast(bands)
    names(r_stack) <- band_names

    # Write output
    param_out_dir <- file.path(output_dir, param)
    dir.create(param_out_dir, recursive = TRUE, showWarnings = FALSE)
    out_path <- file.path(param_out_dir, sprintf("%s_%s.tif", param, tile_id))
    writeRaster(r_stack, out_path, gdal = "COMPRESS=DEFLATE", overwrite = TRUE)

    log_msg("WROTE", sprintf("%s — %d bands", param, nlyr(r_stack)))
  }

  log_msg("DONE", "all params written")
  invisible(tile_id)
}
```

---

## 4. File Discovery

```{r work_units}
if (test_mode) {

  # ── Smoke test: one tile ────────────────────────────────────────────────────
  tile_files   <- list.files(file.path(hist_dir, "A"), pattern = "\\.tif$")
  all_tile_ids <- sub("^A_(.+)\\.tif$", "\\1", tile_files)

  work_units <- list(list(tile_id = all_tile_ids[1]))
  cat(sprintf("[TEST MODE] 1 tile, %d×%d px crop\n", test_crop_n, test_crop_n))

} else {

  # ── Production: one work unit per tile ──────────────────────────────────────
  tile_files   <- list.files(file.path(hist_dir, "A"), pattern = "\\.tif$")
  all_tile_ids <- sub("^A_(.+)\\.tif$", "\\1", tile_files)

  # Check that future tiles match
  future_files   <- list.files(file.path(future_dir, "A"), pattern = "\\.tif$")
  future_tile_ids <- sub("^A_(.+)\\.tif$", "\\1", future_files)
  common_tiles    <- intersect(all_tile_ids, future_tile_ids)

  if (length(common_tiles) < length(all_tile_ids)) {
    cat(sprintf("WARNING: %d hist tiles, %d future tiles, %d in common\n",
                length(all_tile_ids), length(future_tile_ids), length(common_tiles)))
  }


  # Skip completed tiles — only skip if ALL params (A, B, K) exist and are valid
  work_units <- lapply(common_tiles, function(tid) {
    all_done <- all(vapply(params, function(p) {
      f <- file.path(output_dir, p, sprintf("%s_%s.tif", p, tid))
      if (!file.exists(f)) return(FALSE)
      tryCatch({ r <- rast(f); nlyr(r) == length(band_names) },
               error = function(e) FALSE, warning = function(w) FALSE)
    }, logical(1)))
    if (all_done) return(NULL)
    list(tile_id = tid)
  })
  work_units <- Filter(Negate(is.null), work_units)

  cat(sprintf("Tiles: %d total, %d already complete, %d to process\n",
              length(common_tiles),
              length(common_tiles) - length(work_units),
              length(work_units)))
}
```

---

## 5. Output Directory Setup

```{r output_dirs}
invisible(lapply(file.path(output_dir, params),
                 dir.create, recursive = TRUE, showWarnings = FALSE))

log_file <- file.path(output_dir, "progress_interpolate.log")
file.create(log_file, showWarnings = FALSE)
cat(sprintf("Run started: %s\n", format(Sys.time())),
    file = log_file, append = TRUE)

cat(sprintf("Output: %s\n", output_dir))
cat(sprintf("Log:    %s\n", log_file))
```

---

## 6. Interpolation

```{r interpolate_parallel}
n_workers       <- min(n_cores, length(work_units))
per_worker_mem  <- max(1L, floor(terra_mem_gb / max(n_workers, 1L)))

cat(sprintf("=== Interpolation ===\n"))
cat(sprintf("  Work units:   %d\n", length(work_units)))
cat(sprintf("  Workers:      %d (of %d cores)\n", n_workers, parallel::detectCores()))
cat(sprintf("  Per-worker terra memmax: %d GB\n", per_worker_mem))
cat(sprintf("  Started: %s\n", format(Sys.time())))

if (length(work_units) == 0) {
  cat("Nothing to do — all tiles already complete.\n")
} else {
  results <- parallel::mclapply(work_units, function(wu) {
    library(terra)
    terraOptions(memmax = per_worker_mem)
    tryCatch(
      interpolate_tile(
        tile_id     = wu$tile_id,
        hist_dir    = hist_dir,
        future_dir  = future_dir,
        output_dir  = output_dir,
        params      = params,
        weights     = weights,
        band_names  = band_names,
        log_file    = log_file,
        test_crop_n = if (test_mode) test_crop_n else 0L
      ),
      error = function(e) {
        msg <- sprintf("[%s | pid=%d] FATAL: %s\n",
                       wu$tile_id, Sys.getpid(), conditionMessage(e))
        cat(msg, file = log_file, append = TRUE)
        paste("ERROR:", conditionMessage(e))
      }
    )
  }, mc.cores = n_workers)

  # ── Report ──────────────────────────────────────────────────────────────────
  is_err <- vapply(results, function(x) {
    is.character(x) && startsWith(x, "ERROR:")
  }, logical(1))
  cat(sprintf("\n=== Done | %d tiles | %d errors | %s ===\n",
              length(results), sum(is_err), format(Sys.time())))

  if (any(is_err)) {
    cat("\nErrors:\n")
    invisible(lapply(results[is_err], function(e) cat("  ", e, "\n")))
  }
}
```

---

## 7. Run Summary

```{r summary}
for (param in params) {
  n_files <- length(list.files(file.path(output_dir, param), pattern = "\\.tif$"))
  cat(sprintf("  %s: %d tiles\n", param, n_files))
}

if (file.exists(log_file)) {
  cat("\nLast 10 log lines:\n")
  cat(tail(readLines(log_file), 10), sep = "\n")
}
```

---

## 8. Mosaic Tiles into Global Rasters

Merge the 18 per-tile 10-band GeoTIFFs into one global raster per parameter.
Uses `terra::vrt()` to build a GDAL Virtual Raster (lightweight XML pointer)
rather than loading all tiles into memory. `writeRaster()` then streams from
the VRT to disk in chunks, keeping RAM usage low even for large global mosaics.

```{r mosaic}
if (test_mode) {
  message("Mosaic skipped in test mode.")
} else {
  mosaic_dir <- file.path(output_dir, "mosaic")
  dir.create(mosaic_dir, recursive = TRUE, showWarnings = FALSE)

  for (param in params) {
    tile_files <- list.files(file.path(output_dir, param),
                             pattern = "\\.tif$", full.names = TRUE)
    if (length(tile_files) == 0) {
      cat(sprintf("  %s: no tiles found, skipping\n", param))
      next
    }

    # Validate tiles — drop any that GDAL can't read
    valid <- vapply(tile_files, function(tf) {
      tryCatch({ rast(tf); TRUE },
               error = function(e) FALSE, warning = function(w) FALSE)
    }, logical(1))
    if (any(!valid)) {
      cat(sprintf("  %s: dropping %d corrupt tile(s): %s\n",
                  param, sum(!valid),
                  paste(basename(tile_files[!valid]), collapse = ", ")))
      tile_files <- tile_files[valid]
    }

    cat(sprintf("  %s: building VRT from %d tiles ... ", param, length(tile_files)))
    t0 <- Sys.time()

    vrt_path <- file.path(mosaic_dir, sprintf("%s_global.vrt", param))
    v <- vrt(tile_files, filename = vrt_path, overwrite = TRUE)
    names(v) <- band_names

    out_path <- file.path(mosaic_dir, sprintf("%s_global.tif", param))
    writeRaster(v, out_path, gdal = "COMPRESS=DEFLATE", overwrite = TRUE)

    elapsed <- round(difftime(Sys.time(), t0, units = "secs"), 1)
    cat(sprintf("done (%s sec) -> %s\n", elapsed, basename(out_path)))
  }

  cat(sprintf("\nMosaic outputs: %s\n", mosaic_dir))
}
```

---

## 9. Sync Outputs to S3

```{r sync-to-s3}
if (!save_to_s3) {
  message("save_to_s3 is FALSE — skipping S3 sync.")
} else {
  if (!requireNamespace("aws.s3", quietly = TRUE)) install.packages("aws.s3")
  library(aws.s3)

  files <- list.files(output_dir, pattern = "\\.tif$",
                      recursive = TRUE, full.names = TRUE)
  cat(sprintf("Syncing interpolated: %d files -> S3\n", length(files)))

  for (f in files) {
    s3_key <- paste0("serge-wiltshire/nat_for_regen_c_accumulation_data/",
                     "interpolated/", sub(paste0(output_dir, "/"), "", f))
    # Use multipart upload for files > 2 GB to avoid curl integer overflow
    use_multipart <- file.size(f) > 2e9
    tryCatch({
      put_object(file = f, bucket = "sagemaker-gst-stage.sharing",
                 object = s3_key, multipart = use_multipart)
    }, error = function(e) {
      cat(sprintf("  Failed: %s — %s\n", basename(f), conditionMessage(e)))
    })
  }
  cat("Done.\n")
}
```

---

*Interpolation completes in seconds per tile — much faster than the CR fitting step.*
