---
title: "Input Data Acquisition"
subtitle: "GCS to S3 (one-time) and S3 to SageMaker (each instance)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

---

## Overview

Two-step data pipeline:

1. **GCS to S3** — run **once** to pay GCS egress once and land data durably
   in S3.
2. **S3 to SageMaker local** — run at the **start of each SageMaker session**
   to pull data from S3 to local instance storage for fast parallel I/O during
   the CR fitting run.

All shell commands run in the RStudio **Terminal tab** (not the Console).
`rclone` is used throughout — the AWS CLI is not available in the SageMaker
Posit shell.

---

## 1. Configuration

```{r config}
# GCS source ------------------------------------------------------------------
gcs_bucket <- "carbon-accumulation-modeling-bucket"
gcs_prefix <- "rf_outputs"

# S3 --------------------------------------------------------------------------
s3_bucket  <- "sagemaker-gst-stage.sharing"
s3_prefix  <- "serge-wiltshire/nat_for_regen_c_accumulation_data"

# Scenarios -------------------------------------------------------------------
scenarios  <- c("future", "hist")

# Local paths (SageMaker) -----------------------------------------------------
local_repo   <- "/home/sagemaker-user/nat_for_regen_c_accumulation_sw"
local_inputs <- file.path(local_repo, "data", "inputs")
```

---

## 2. GCS to S3 (one-time transfer)

> Run this section **once** from the SageMaker Terminal. Data stays in S3
> permanently — re-running is safe (`rclone copy` skips existing files).

### 2.1 Install rclone

```{bash install-rclone}
curl https://rclone.org/install.sh | sudo bash
rclone --version   # confirm install
```

### 2.2 Create a GCP service account

rclone needs credentials to read from GCS. A service account JSON key is the
right approach for a server environment without a browser.

1. Go to [console.cloud.google.com](https://console.cloud.google.com)
2. Hamburger menu (≡) → **IAM & Admin** → **Service Accounts**
3. **Create Service Account** → name it `rclone-gcs-reader` → **Create and Continue**
4. Grant role: **Storage Object Viewer** → **Continue** → **Done**
5. Click the new service account → **Keys** tab → **Add Key** → **Create new key** → **JSON** → **Create** (downloads automatically)
6. Go to **Cloud Storage** → **Buckets** → click `carbon-accumulation-modeling-bucket` → **Permissions** → **Grant Access** → paste the service account email → role: **Storage Object Viewer** → **Save**
7. Upload the JSON key to SageMaker via the RStudio **Files pane → Upload**

### 2.3 Configure rclone remotes

Run `rclone config` in the terminal and follow the prompts. Create two remotes
using the settings below.

```{bash rclone-config}
rclone config
```

#### GCS remote

| Prompt | Enter | Note |
|--------|-------|------|
| `n` for new remote | `n` | |
| name | `gcs` | Used in all transfer commands |
| Storage type | `Google Cloud Storage` | Search or select number |
| client_id | Enter | Skip |
| client_secret | Enter | Skip |
| project_number | Enter | Skip — only needed for creating buckets |
| user_project | Enter | Skip |
| service_account_file | `/home/sagemaker-user/<key-filename>.json` | Full path to uploaded JSON key |
| anonymous | Enter | Skip (default: false) |
| object_acl | Enter | Skip — read-only use |
| bucket_acl | Enter | Skip — read-only use |
| bucket_policy_only | Enter | Skip |
| location | Enter | Skip |
| storage_class | Enter | Skip |
| env_auth | Enter | Skip (default: false — using service account file) |
| Edit advanced config | Enter | Skip (default: no) |

#### S3 remote

The SageMaker instance has an IAM role attached — `env_auth = true` picks up
credentials automatically; no access key or secret needed.

Find your region first:

```{bash find-region}
echo $AWS_DEFAULT_REGION   # returns e.g. us-west-2
```

| Prompt | Enter | Note |
|--------|-------|------|
| `n` for new remote | `n` | |
| name | `s3aws` | Used in all transfer commands |
| Storage type | `Amazon S3` | Search or select number |
| Provider | `AWS` | |
| env_auth | `true` | Uses SageMaker IAM role automatically |
| access_key_id | Enter | Skip — env_auth handles credentials |
| secret_access_key | Enter | Skip — env_auth handles credentials |
| region | `us-west-2` | Match your SageMaker instance region |
| endpoint | Enter | Skip — standard AWS S3 |
| location_constraint | Enter | Skip |
| acl | Enter | Skip |
| server_side_encryption | Enter | Skip |
| sse_kms_key_id | Enter | Skip |
| storage_class | Enter | Skip |
| bucket_object_lock_enabled | Enter | Skip |
| Edit advanced config | Enter | Skip (default: no) |

### 2.4 Verify remotes

Confirm both remotes can see what they need before starting the transfer.

```{bash verify-remotes}
# Should show future/ and hist/ directories
rclone lsd gcs:carbon-accumulation-modeling-bucket/rf_outputs/

# Should show the serge-wiltshire/ prefix (may be empty before first transfer)
rclone lsd s3aws:sagemaker-gst-stage.sharing/serge-wiltshire/
```

### 2.5 Transfer data

Run each scenario separately. Use `nohup` so the transfer survives a session
disconnect, and `--log-file` to monitor progress.

```{bash transfer-future}
nohup rclone copy \
  "gcs:carbon-accumulation-modeling-bucket/rf_outputs/future/" \
  "s3aws:sagemaker-gst-stage.sharing/serge-wiltshire/nat_for_regen_c_accumulation_data/future/inputs/" \
  --transfers 20 \
  --stats 10s \
  --log-file ~/rclone_future.log &

tail -f ~/rclone_future.log
```

`Ctrl+C` stops the log tail — the transfer keeps running in the background.

```{bash transfer-hist}
nohup rclone copy \
  "gcs:carbon-accumulation-modeling-bucket/rf_outputs/hist/" \
  "s3aws:sagemaker-gst-stage.sharing/serge-wiltshire/nat_for_regen_c_accumulation_data/hist/inputs/" \
  --transfers 20 \
  --stats 10s \
  --log-file ~/rclone_hist.log &

tail -f ~/rclone_hist.log
```

### 2.6 Verify transfer

Confirm file counts match between GCS and S3.

```{bash verify-transfer}
for SCENARIO in future hist; do
  N_GCS=$(rclone ls "gcs:carbon-accumulation-modeling-bucket/rf_outputs/${SCENARIO}/"                                                           | wc -l)
  N_S3=$(rclone  ls "s3aws:sagemaker-gst-stage.sharing/serge-wiltshire/nat_for_regen_c_accumulation_data/${SCENARIO}/inputs/" | wc -l)
  echo "${SCENARIO} -- GCS: ${N_GCS} files | S3: ${N_S3} files"
done
```

---

## 3. S3 to SageMaker local (run at instance start)

> Run this section each time you start a new SageMaker session.
> `rclone copy` skips files that already exist, so re-running is fast.

### 3.1 Clone the repo (first time only)

```{bash clone-repo}
cd /home/sagemaker-user
git clone https://github.com/swiltshire/nat_for_regen_c_accumulation_sw.git
```

If the repo is private, use your GitHub username and a Personal Access Token
(not your password) when prompted. Generate one at:
GitHub → Settings → Developer settings → Personal access tokens → Generate new
token (scope: `repo`).

### 3.2 Sync inputs from S3

```{bash sync-future}
rclone copy \
  "s3aws:sagemaker-gst-stage.sharing/serge-wiltshire/nat_for_regen_c_accumulation_data/future/inputs/" \
  "/home/sagemaker-user/nat_for_regen_c_accumulation_sw/data/inputs/future/" \
  --transfers 20 --progress
```

```{bash sync-hist}
rclone copy \
  "s3aws:sagemaker-gst-stage.sharing/serge-wiltshire/nat_for_regen_c_accumulation_data/hist/inputs/" \
  "/home/sagemaker-user/nat_for_regen_c_accumulation_sw/data/inputs/hist/" \
  --transfers 20 --progress
```

### 3.3 Verify

```{r verify-local}
# Expected: 360 files per scenario (20 ages x 18 tiles)
for (scenario in scenarios) {
  d <- file.path(local_inputs, scenario)
  n <- length(list.files(d, pattern = "\\.tif$"))
  status <- if (n == 360) "OK" else sprintf("WARNING: expected 360, got %d", n)
  cat(sprintf("%-8s  %d files  [%s]\n", scenario, n, status))
}
```
