---
title: "Input Data Acquisition"
subtitle: "GCS to S3 (one-time) and S3 to SageMaker (each instance)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

---

## Overview

Two-step data pipeline:

1. **GCS to S3** - run **once** from a local machine (or any environment with
   `rclone` and credentials for both GCS and S3). Pays GCS egress once; data
   then lives durably in S3.
2. **S3 to SageMaker** - run at the **start of each SageMaker session** to sync
   data from S3 to the instance's local storage for fast parallel I/O during
   the CR fitting run.

---

## 1. Configuration

Set these variables once - they are referenced throughout the notebook.

```{r config}
# GCS source -----------------------------------------------------------------
# Verify the exact folder structure in the bucket before running Section 2.
gcs_bucket <- "carbon-accumulation-modeling-bucket"
gcs_prefix <- "rf_outputs"           # path inside bucket to the scenario dirs

# S3 destination -------------------------------------------------------------
s3_bucket  <- "your-s3-bucket-name"  # replace with your bucket
s3_prefix  <- "nat_for_regen/data/inputs"

# Scenarios ------------------------------------------------------------------
scenarios  <- c("future", "hist")

# Local path (SageMaker) -----------------------------------------------------
# Absolute path to the cloned repo on the SageMaker instance.
local_repo   <- "/home/ec2-user/SageMaker/nat_for_regen_c_accumulation_sw"
local_inputs <- file.path(local_repo, "data", "inputs")
```

---

## 2. GCS to S3 (one-time transfer)

> **Run this section locally** (or on any machine with `rclone` installed and
> credentials for both GCS and S3). Do **not** run on SageMaker - GCS egress
> fees are incurred regardless of origin, and running it here saves nothing.

### 2.1 Install rclone

```{bash install-rclone}
# macOS
brew install rclone

# Linux
curl https://rclone.org/install.sh | sudo bash
```

### 2.2 Configure remotes

Run `rclone config` interactively to create two named remotes.

```{bash rclone-config}
rclone config
```

**GCS remote** (name it `gcs`):

- Type: `Google Cloud Storage`
- Auth: `false` - uses application default credentials
- Run `gcloud auth application-default login` beforehand, or supply a service
  account JSON key path when prompted.

**S3 remote** (name it `s3aws`):

- Type: `Amazon S3`
- Provider: `AWS`
- Auth: enter your AWS access key and secret, **or** leave blank if your
  environment inherits credentials from `~/.aws/credentials`.
- Region: match your S3 bucket region (e.g. `us-east-1`)

Verify both remotes before transferring:

```{bash verify-remotes}
rclone lsd gcs:carbon-accumulation-modeling-bucket/rf_outputs/
rclone lsd s3aws:your-s3-bucket-name/nat_for_regen/data/inputs/
```

### 2.3 Transfer

`--transfers 20` parallelises file-level transfers. Lower this if you see
connection errors on a slower link.

```{bash gcs-to-s3}
GCS_BUCKET="carbon-accumulation-modeling-bucket"
GCS_PREFIX="rf_outputs"
S3_BUCKET="your-s3-bucket-name"
S3_PREFIX="nat_for_regen/data/inputs"

for SCENARIO in future hist; do
  echo "=== Transferring ${SCENARIO} ==="
  rclone copy \
    "gcs:${GCS_BUCKET}/${GCS_PREFIX}/${SCENARIO}/" \
    "s3aws:${S3_BUCKET}/${S3_PREFIX}/${SCENARIO}/" \
    --transfers 20 \
    --progress \
    --stats-one-line
done
```

Verify file counts match the source:

```{bash verify-s3}
GCS_BUCKET="carbon-accumulation-modeling-bucket"
GCS_PREFIX="rf_outputs"
S3_BUCKET="your-s3-bucket-name"
S3_PREFIX="nat_for_regen/data/inputs"

for SCENARIO in future hist; do
  N_GCS=$(rclone ls "gcs:${GCS_BUCKET}/${GCS_PREFIX}/${SCENARIO}/" | wc -l)
  N_S3=$(rclone  ls "s3aws:${S3_BUCKET}/${S3_PREFIX}/${SCENARIO}/" | wc -l)
  echo "${SCENARIO} -- GCS: ${N_GCS} files | S3: ${N_S3} files"
done
```

---

## 3. S3 to SageMaker (run at instance start)

> **Run this section on SageMaker.** The instance IAM role must have
> `s3:GetObject` and `s3:ListBucket` on your bucket - no explicit credentials
> are needed if the role is configured correctly.

### 3.1 Clone the repo (first time only)

```{bash clone-repo}
cd /home/ec2-user/SageMaker
git clone git@github.com:swiltshire/nat_for_regen_c_accumulation_sw.git
```

### 3.2 Sync inputs from S3

`aws s3 sync` skips files that already exist and match, so re-running at the
start of each session is safe and fast when nothing has changed.

```{bash s3-to-sagemaker}
S3_BUCKET="your-s3-bucket-name"
S3_PREFIX="nat_for_regen/data/inputs"
LOCAL_INPUTS="/home/ec2-user/SageMaker/nat_for_regen_c_accumulation_sw/data/inputs"

mkdir -p "${LOCAL_INPUTS}"

for SCENARIO in future hist; do
  echo "=== Syncing ${SCENARIO} from S3 ==="
  aws s3 sync \
    "s3://${S3_BUCKET}/${S3_PREFIX}/${SCENARIO}/" \
    "${LOCAL_INPUTS}/${SCENARIO}/" \
    --no-progress
  echo "Done: ${SCENARIO}"
done
```

### 3.3 Verify

```{r verify-local}
# Expected: 360 files per scenario (20 ages x 18 tiles)
for (scenario in scenarios) {
  d <- file.path(local_inputs, scenario)
  n <- length(list.files(d, pattern = "\\.tif$"))
  status <- if (n == 360) "OK" else sprintf("WARNING: expected 360, got %d", n)
  cat(sprintf("%-8s  %d files  [%s]\n", scenario, n, status))
}
```

---

## 4. Sync outputs back to S3

Run this after the CR fitting job completes, before terminating the instance.
SageMaker instance storage is ephemeral - outputs not synced to S3 will be lost.

```{bash upload-outputs}
S3_BUCKET="your-s3-bucket-name"
LOCAL_OUTPUTS="/home/ec2-user/SageMaker/nat_for_regen_c_accumulation_sw/data/outputs"

aws s3 sync \
  "${LOCAL_OUTPUTS}/" \
  "s3://${S3_BUCKET}/nat_for_regen/data/outputs/" \
  --no-progress

echo "Outputs synced to S3."
```
